---
title: "testTest-vignette"
output: rmarkdown::html_vignette
author: "Piska46"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{testTest-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(testTest)
```

## Intro

The sole purpuse of this tiny packacge (only 4 lines of code) is to demonstrate 
a difference in verdicts made by 3 kinds of tests and to provide a remedy.

The package used an environment with data that may be manipulated. And, in order to get 
predictable results each test file should start with the same 'factory' data.  However, achieving this
by detaching, unloading and re-requiring the package at the start of each test file will causes a test problem.

Different way of testing rersulted in different verdicts.

## Overview


1. The intended behaviour of this package is demonstrated.

2. The 'setMenu' function is tested with a simple use of 'testthat' This works well.

3. The 'setMenu' function is tested with 'rcmdcheck()' Which works well.

4. The 'setMenu' function is tested with 'devtools' Which caused a problem.

## Demonstration of the testest package

The 'factory set' menu is printed, then it is set to the new menu and the new menu 
should be printed correctly

```{r}
printMenu ()
setMenu ("Bacon")
printMenu ()
```
Conclusion: it works perfectly.

## The test file use for all tests

The package test-file 'testTestTest.R' reads:

```{r testTestTest.R}
require(testthat)

# laoding a fresh package
if (isNamespaceLoaded ("testTest")) detach (package:testTest, unload = TRUE)
require (testTest)
testEnvir <- testTest:::testEnvir

test_that ("testTestTest", {
    expect_equal (testEnvir$menu, "spam")
    testEnvir$menu <- "eggs"
    expect_equal (testEnvir$menu, "eggs")

    setMenu ("bacon")
    expect_equal (testEnvir$menu, "bacon")   # fails with devtools::test() only
})
```

Note that the testTest package is detached, unloaded, and re-required at the beginning.
This is to make shure that the test is preformed on testEnvir data with 'factory' settings, i.e. on data
that is not changed with 'setMenu' functions.

However as shown below, this will cause a severe problem.

## A simple test with testthat

This is the output produced using source (".\tests\testthat\testTestTest.R")

    Test passed 
    Loading required package: testTest
    Attaching package: ‘testTest’
    The following object is masked _by_ ‘.GlobalEnv’:
        testEnvir
    Test passed 

This is the output produced when pressing the RStudio top right "Run Test" button:

    ==> Testing R file using 'testthat'
    ℹ Loading testTest
    [ FAIL 0 | WARN 0 | SKIP 0 | PASS 3 ]Loading required package: testTest
    [ FAIL 0 | WARN 0 | SKIP 0 | PASS 4 ]
    Test complete

## Running rcmdcheck::rcmdcheck()

This is the result when running rcmdcheck::rcmdcheck()

── R CMD check results ───────────────────────────────────────────────────────────────────────────────────────────────────── testTest 0.1.0 ────
Duration: 12.8s

❯ checking DESCRIPTION meta-information ... WARNING
  Non-standard license specification:
    What license is it under?
  Standardizable: FALSE

❯ checking for missing documentation entries ... WARNING
  Undocumented code objects:
    ‘printMenu’ ‘setMenu’ ‘testEnvir’
  All user-level objects in a package should have documentation entries.
  See chapter ‘Writing R documentation files’ in the ‘Writing R
  Extensions’ manual.

0 errors ✔ | 2 warnings ✖ | 0 notes ✔
>

NOTE: No complains about the value of testEnvir$menu


## Using devtools

This is the result when running devtools:::test_active_file()

    > devtools:::test_active_file()
    [ FAIL 0 | WARN 0 | SKIP 0 | PASS 0 ]Loading required package: testTest
    
    Attaching package: ‘testTest’
    
    The following object is masked _by_ ‘.GlobalEnv’:
    
        testEnvir
    
    [ FAIL 1 | WARN 0 | SKIP 0 | PASS 2 ]
    
    ── Failure (testTestTest.R:15:5): testTestTest ─────────────────────────────────
    testEnvir$menu (`actual`) not equal to "bacon" (`expected`).
    
    `actual`:   "eggs" 
    `expected`: "bacon"
    [ FAIL 1 | WARN 0 | SKIP 0 | PASS 2 ]


It looks like devtools:::test_active_file() keeps using the 'old' testEnvir. i.e. the one
that exists when the function in invoked. And, that 'setMenu ("bacon")' used - as required for this package - 
the new environment made available after re-requiring the 'testTest' package.

## Concluding considerations

As devtools:::test_active_file() may produce a negative verdict for expect_... cases may be ok for the package at hand
and for rcmdcheck or sourcing a test file, it may impose unnecessary drawbacks for some package developers. 
This holds especially in case the package contains data that may extended by users who may e.g. add new 
objects and/or add/modify object attributes and/or set/modify user preferences (e.g. for specific objects).

However there does exists a simple work around.

## A simple remedy

Make a copy of all vulnerable 'factory' data in the environment under test and restore it at the 
beginning of each test file.

Add, e.g. a function like this in "testthat/helper.R"

    resetMyData <- function () {
        if (myBackup doesn't yet extsts) {
            backup vulnarable data  # will be excecuted for the first testfile only
        } else {
            restore from backup
        }
    }

And add resetMyData() to the beginning of each test file.
